---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Migration data



This repo houses an in-progress piece of data visualisation/journalism/art related to the U.S. Census data on migration. It uses 2018 migration data to explore annual trends in where Americans are moving within the country. The original data was downloaded here:

https://www.census.gov/data/tables/time-series/demo/geographic-mobility/state-to-state-migration.html


# Maps

The repo contains a base script with a few various leaflet maps created in the reporting, e.g.:

```{r map, echo = FALSE, warning = FALSE, message = FALSE}
library(openxlsx)
library(dplyr)

migration <- openxlsx::read.xlsx("scrolly/data/migration_flow_clean.xlsx")

mig2 <- migration[,-c(2:9)]
mig3 <- mig2[,c(FALSE, TRUE)]

colnames(mig3) <- mig3[1,]

mig4 <- mig3[-c(1,2),]

states <- as.data.frame(migration[-c(1,2),1])

colnames(states) <- "destinations"

migration2 <- cbind(states, mig4)

mig_long <- migration2 %>%
  tidyr::pivot_longer(-destinations, names_to = "origins", values_to = "count")

mig_un <- mig_long[-c(1:55),]

mig_states <- mig_un %>%
  dplyr::select(origins, destinations, count)


library(leaflet)
library(RColorBrewer)

#library(rnaturalearth) # devtools::install_github('ropenscilabs/rnaturalearth')

# countries <- ne_countries()

#states <- ne_states(iso_a2 = 'US')

# states_xy <- states@data %>%
#   select(name, longitude, latitude)


load(file = "scrolly/data/states_xy.rda")

library(rgdal)

mig_lat <- mig_states %>%
  left_join(states_xy, by = c('origins' = 'name')) %>%
  left_join(states_xy, by = c('destinations' = 'name'))

mig_clean <- mig_lat %>%
  dplyr::filter(!is.na(longitude.x),
                !is.na(longitude.y),
                !count=="N/A",
                !count==0)

mig_weight <- mig_clean %>%
  mutate(weight = scales::rescale(as.numeric(count), to = c(0.02, 10)),
         count = as.numeric(count),
         origins = as.factor(origins),
         destinations = as.factor(destinations))

mig_ori <- mig_weight %>%
  group_by(origins) %>%
  summarise(state_ori = sum(count))

mig_dest <- mig_weight %>%
  group_by(destinations) %>%
  summarise(state_dest = sum(count))

mig_diff <- left_join(mig_dest, mig_ori, by=c("destinations" = "origins")) %>%
  mutate(diff = abs(state_ori - state_dest))

mig_weight_top <- mig_weight %>%
  dplyr::filter(weight > 2.19)

mig_weight_bot <- mig_weight %>%
  dplyr::filter(weight <= 2.19)



library(geosphere)

flows <- gcIntermediate(mig_weight[,4:5], mig_weight[,6:7], sp = TRUE, addStartEnd = TRUE)



flows$counts <- mig_weight$count
flows$origins <- mig_weight$origins
flows$destinations <- mig_weight$destinations
flows$weight <- mig_weight$weight

hover <- paste0(flows$origins, " to ", 
                flows$destinations, ': ', 
                as.character(flows$counts))



flows_top <- gcIntermediate(mig_weight_top[,4:5], mig_weight_top[,6:7], sp = TRUE, addStartEnd = TRUE)

flows_top$counts <- mig_weight_top$count
flows_top$origins <- mig_weight_top$origins
flows_top$destinations <- mig_weight_top$destinations
flows_top$weight <- mig_weight_top$weight

flows_bot <- gcIntermediate(mig_weight_bot[,4:5], mig_weight_bot[,6:7], sp = TRUE, addStartEnd = TRUE)

flows_bot$counts <- mig_weight_bot$count
flows_bot$origins <- mig_weight_bot$origins
flows_bot$destinations <- mig_weight_bot$destinations
flows_bot$weight <- mig_weight_bot$weight


hover_top <- paste0(flows_top$origins, " to ", 
                    flows_top$destinations, ': ', 
                    as.character(flows_top$counts))

hover_bot <- paste0(flows_bot$origins, " to ", 
                    flows_bot$destinations, ': ', 
                    as.character(flows_bot$counts))

pal <- colorFactor(brewer.pal(4, 'Set2'), flows$origins)

library(shiny)

tag.map.title <- tags$style(HTML("
  .leaflet-control.map-title { 
    transform: translate(-50%,20%);
    position: fixed !important;
    left: 50%;
    padding-left: 10px; 
    padding-right: 10px; 
    background: rgba(255,255,255,0.75);
    font-weight: bold;
    font-size: 26px;
  }
"))

title <- tags$div(
  tag.map.title, HTML("Most Common Migrations in the U.S. in 2018")
)  


leaflet() %>%
  addProviderTiles('CartoDB.Positron') %>%
  addPolylines(data = flows_bot, weight = ~weight,
               fillOpacity = .2,
               group = ~origins, color = "#d3d3d3") %>%
  addPolylines(data = flows_top, label = hover_top, 
               labelOptions = labelOptions(
                 style = list("font-weight" = "normal", padding = "3px 8px"),
                 textsize = "15px",
                 direction = "auto"),
               weight = ~weight,
               opacity = .6,
               group = ~origins, color = "#1167b1",
               highlight = highlightOptions(opacity = 0.8, color = "red", bringToFront = TRUE)) %>%
  setView(-98.5795, 39.8283, zoom = 3.5) %>%
  addControl(title, position = "topleft", className="map-title")
  
```

# Scrolly telling

The `scrolly` subdirectory has a shiny application displaying the the piece of scrolly telling. Thhe functionality to detect scroll by the user and update the map based on scroll was taken from a John Coene blogpost on scrollytelling in R. The repo with tthat blogpost and the dependencies for this to be reproducible are found here:

https://github.com/JohnCoene/scrollytell

The final product is here:

https://willdebras.shinyapps.io/migration_2018/
